---
title: "GIS Final Project"
author: "Homar A. Maurás Rodríguez & Tianyu Bai"
format: html
editor: visual
execute: 
  echo: false
  error: false
---

```{r libraries}
library(tidyverse) |> suppressPackageStartupMessages()
library(dplyr) |> suppressPackageStartupMessages()
library(sf) |> suppressPackageStartupMessages()
library(leaflet) |> suppressPackageStartupMessages()
library(mapview) |> suppressPackageStartupMessages()
library(lubridate) |> suppressPackageStartupMessages()
```

```{r datasets}
acled <- read.csv('/Users/homi/Data Science/datasets/acled_aggregated_20060.csv')
confessional <- read_csv('/Users/homi/Data Science/datasets/adm1_confessional.csv')
lebanon_adm1 <- st_read("/Users/homi/Data Science/datasets/geoBoundaries-LBN-ADM1-all/geoBoundaries-LBN-ADM1.shp")
```

```{r formatting}
#lowercase
names(acled) <- tolower(names(acled))
names(confessional) <- tolower(names(confessional))
names(lebanon_adm1) <- tolower(names(lebanon_adm1))

#selecting relevant variables
acled <- acled %>% select(admin1, event.month, sub.event.type, events)
lebanon_adm1 <- lebanon_adm1 %>% select(shapename, shapeiso, geometry)

#renaming variables
acled <- acled %>% rename(adm1 = admin1, event_month = event.month, event_type = sub.event.type)
confessional <- confessional %>% rename(iso = iso_3166_2, sect = reported_sect, total_sect = adm1_total)
lebanon_adm1 <- lebanon_adm1 %>% rename(adm1 = shapename, iso = shapeiso)

#replace event_month dates with corresponding month name (using lubridate)
acled$event_month <- ymd(acled$event_month) #converting event_month into date format
acled$prior_month <- floor_date(acled$event_month - days(1), "month") #adjust to the prior month by subtracting one day, then getting the first day of that month
acled$prior_month <- format(acled$prior_month, "%B, %Y") #format the 'prior_month' column as "Month, Year"
acled <- acled %>% relocate(prior_month, .before = event_month) #changing column order
```

```{r names}
# replace sect to Maronite, Christian Minorities, Shia, Sunni, Druze, and Other
confessional <- confessional %>%
  mutate(sect = if_else(sect %in% c("Armenian Catholic", "Armenian Orthodox", 
                                    "Christian Minorities", "Greek Catholic", 
                                    "Greek Orthodox", "Protestant"), 
                        "Christian Minorities", sect)) #replacing christian groups into broader christian category
confessional <- confessional %>%
  mutate(sect = if_else(sect %in% "Alawite", 
                        "Shia", sect)) #including Alawite in Shia faith

#replacing adm1 names to english
confessional <- confessional %>% #changing confessional
  mutate(adm1 = if_else(adm1 %in% "Bayrut", "Beirut", adm1),
         adm1 = if_else(adm1 %in% "Beqaa", "Bekaa", adm1),
         adm1 = if_else(adm1 %in% "Liban-Nord", "North", adm1),
         adm1 = if_else(adm1 %in% "Liban-Sud", "South", adm1),
         adm1 = if_else(adm1 %in% "Mont-Liban", "Mount Lebanon", adm1),
         adm1 = if_else(adm1 %in% "Nabatiye", "Al Nabatieh", adm1)) 

lebanon_adm1 <- lebanon_adm1 %>%
  mutate(adm1 = if_else(adm1 %in% "Aakkâr", "Akkar", adm1),
         adm1 = if_else(adm1 %in% "Nabatîyé", "Al Nabatieh", adm1),
         adm1 = if_else(adm1 %in% "Beyrouth", "Beirut", adm1),
         adm1 = if_else(adm1 %in% "Béqaa", "Bekaa", adm1),
         adm1 = if_else(adm1 %in% "Liban-Nord", "North", adm1),
         adm1 = if_else(adm1 %in% "Mont-Liban", "Mount Lebanon", adm1),
         adm1 = if_else(adm1 %in% "Liban-Sud", "South", adm1),
         adm1 = if_else(adm1 %in% "Nabatîyé", "Al Nabatieh", adm1))
```

# Understanding Israeli attack patterns across ethno-religious groups in Lebanon

## I. Introduction

The history of conflict between Israel and Lebanon, particularly involving Hezbollah, is marked by recurring hostilities and complex ethno-religious dynamics. The most recent escalation began on October 7, 2023 which led to a 14-month conflict between Israel and Hezbollah. The recent escalation between Israel and Hezbollah has had profound implications for Lebanon’s ethno-religious groups. Shiite communities have borne the brunt of the conflict due to their association with Hezbollah, while some Christian areas have remained relatively unscathed. Understanding these patterns is crucial for analyzing the broader socio-political impact of the conflict on Lebanon's diverse society.

### 1.1 Objective

This study aims to explore the the impact of conflict dynamics on ethno-religious identities in Lebanon, focusing on how Israeli attack patterns disproportionately impact different groups and regions. By examining spatial and identity-based disparities in attack frequency, types, and casualties, the research seeks to uncover trends that can inform humanitarian aid, guide policy interventions, and the visibilization of those who bare the cost of war.

#### 1.1.1 Research Questions

1.  What is the likelihood of an attack occurring in a region based on the predominant ethno-religious identity?
2.  Do casualty rates vary across ethno-religious groups?
3.  Are specific types of attacks more prevalent in areas dominated by particular identity groups?

#### 1.1.2 Variables of Interest

-   Identity: Captures the impact of attacks on various ethno-religious groups, highlighting potential disparities in targeting or vulnerability.
-   Governorates: Analyzes spatial trends in attacks by administrative regions, providing a geographic perspective on conflict dynamics.
-   Attack type: Identifies the nature of the attacks to assess whether certain tactics are disproportionately used against specific groups or areas.
-   Casualties: Examines the human cost of conflict, measuring variations in the number of casualties across regions and identity groups.

#### 1.1.3 Significance

Understanding the spatial distribution of attacks across identity groups is critical for shaping effective humanitarian aid interventions and informing conflict mitigation strategies. By identifying patterns, this research can support targeted resource allocation and long-term reparations.

## II. Data and Methodology

### 2.1 Data Sources

-   Population statistics: World Health Organization (WHO) Data & CIA World Handbook
-   Confessional distribution: United Nations Office for the Coordination of Humanitarian Affairs (UN OCHA)
-   Administrative data: Human Data Exchange
-   Attack reports: Armed Conflict Location & Event Data (ACLED)

### 2.2 Limitations and Assumptions

#### 2.2.1 Limitations

-   Despite the sizeable refugee population in Lebanon, due to time constraints and lack of readily available data on the said groups, they will not be included in the analysis.
-   The absence of point data for attack types limits analysis to aggregating at the administrative 1 level.
-   The last census in Lebanon was performed in 1932. The Absence of publicly accessible and recurring census data limits the depth of the analysis.

#### 2.2.2 Assumptions

-   Rates of confessional distribution are constant.
-   Attack patterns affect identity groups at the same rate within a same Governorate

### 2.3 Scope

#### 2.3.1 Level of Analysis

The variables of interest are measured at the Governorate (administrative 1) level. Lebanon is divided into 9 Governorates:

1.  Akkar
2.  Baalbeck-Hermel
3.  Beirut (capital)
4.  Bekaa
5.  Mount Lebanon
6.  Al Nabatieh
7.  North Lebanon
8.  South Lebanon
9.  Keserwan-Jbeil

```{r og_map}
adm1_map <- lebanon_adm1 %>%
  select(adm1, geometry) %>%
  mapview(layer.name = "Governorate")
adm1_map
```

Keserwan-Jbeil was officially established in 2017, separating from Mount Lebanon; thus, the available data only includes estimates for eight out of the nine governorates.

```{r union}
#combining Jbiel and Mount Lebanon
lebanon_union <- lebanon_adm1 %>%
  slice(c(4,9)) %>% 
  mutate(geometry = st_make_valid(geometry)) %>%
  summarize(geometry = st_union(geometry)) %>%
  mutate(adm1 = "Mount Lebanon") %>%
  mutate(iso = "LB-JL")
  
lebanon_adm1 <- lebanon_adm1 %>%
  slice(-c(4,9)) %>% 
  bind_rows(lebanon_union)

union_map <- lebanon_adm1 %>%
  select(adm1, geometry) %>%
  mapview(layer.name = "Governorate")
union_map 
```

### 2.4 Demographics

As stated previously, the last census in Lebanon was performed in 1932– consequently, the absence of transparent data on the country’s demographic makeup becomes the largest part of the puzzle. As of 2023, the World Health Organization (WHO) estimates Lebanon's population at approximately 5.77 million. Recent U.S. State Department reports indicate that Syrian refugees account for around 1.5 million of this population, while Palestinian refugees are estimated at 250,000. Other refugee groups collectively number, approximately, 11,645. While there are no official figures from the Lebanese government, our closest proxy for population is the number of registered voters, which is:

```{r voters}
sum(confessional$total_sect)
```


#### 2.4.1 Confessional Distribution in Lebanon

Lebanon’s political system is based on a confessional model, where power is divided among these groups:

1.  Maronites
2.  Non-Maronite Christian Minorities
3.  Sunni Muslims
4.  Shia Muslims
5.  Druze

This arrangement has fostered cooperation but also entrenched sectarian divisions, influencing everything from governance to social interactions.

*Maronites*

Maronites are Lebanon's largest Christian sect and hold a unique position in the country's socio-political landscape. In communion with the Roman Catholic Church, the Maronites have maintained distinct liturgical practices, contributing to their strong cultural identity. Historically, they have been one of the most politically powerful groups in Lebanon, with the presidency reserved for a Maronite under the National Pact of 1943. This arrangement solidified their influence during Lebanon's formative years, allowing them to shape the nation's political and economic systems. Their communities are predominantly concentrated in Mount Lebanon and parts of Beirut, where they have historically formed the backbone of Lebanon's middle and upper classes. Despite demographic changes and shifting political dynamics, the Maronites continue to wield significant influence, often serving as a bridge between Lebanon's Christian and Muslim populations.

*Non-Maronite Christian Minorities* 

Non-Maronite Christian minorities in Lebanon encompass a diverse range of denominations, including Greek Orthodox, Greek Catholics (Melkites), Armenian Orthodox, Armenian Catholics, Protestants, and other smaller Christian groups. Unlike the Maronites, these groups have not historically enjoyed the same level of political power or visibility. However, they play a crucial role in Lebanon's religious mosaic, often serving as cultural and economic contributors within their communities. The Greek Orthodox, for instance, traditionally hold the position of Deputy Prime Minister, reflecting their role in Lebanon's sectarian power-sharing arrangement. Armenian Christians, concentrated in areas like Bourj Hammoud and Anjar, have cultivated strong communal networks and cultural institutions, helping to preserve their heritage. While non-Maronite Christian minorities often align with the broader Christian political agenda, their distinct identities and historical experiences set them apart within Lebanon's complex confessional landscape.

*Sunni Muslim* 

Sunni Muslims, who adhere to the Sunni branch of Islam, form one of Lebanon's key religious communities and hold a historically prominent role in the country's socio-political framework. Sunnis emphasize adherence to the traditions of the Prophet Muhammad, as interpreted through scholarly consensus, making them part of the broader Sunni majority in the Islamic world. In Lebanon's sectarian power-sharing system, the position of Prime Minister is reserved for a Sunni Muslim, underscoring their political significance. Sunni communities are predominantly located in major urban centers such as Beirut, Tripoli, and Sidon, as well as parts of the Bekaa Valley and southern Lebanon. Historically, they have been influential in Lebanon's economic and political spheres, often acting as a counterbalance to Shia and Christian groups. Despite internal divisions and the challenges posed by regional and domestic conflicts, Sunnis continue to play a vital role in shaping Lebanon’s political discourse and development.

*Shia Muslims* 

Shia Muslims, the largest Muslim sect in Lebanon, follow the Shia branch of Islam, which emphasizes the spiritual and political leadership of the Prophet Muhammad's descendants, particularly Ali and his lineage. Traditionally marginalized within Lebanon's political and economic systems, the Shia community experienced a significant transformation in the late 20th century with the emergence of Hezbollah. Established in the 1980s, Hezbollah has become a dominant political and military force, providing the Shia community with unprecedented political representation and a voice in regional geopolitics. Shia populations are concentrated in southern Lebanon, the Bekaa Valley, and the southern suburbs of Beirut (known as Dahiyeh), areas that have faced disproportionate challenges during periods of conflict. Their association with Hezbollah has made them a focal point of political contention and military targeting, positioning them at the center of Lebanon's contemporary struggles. Despite these challenges, the Shia community remains a crucial and dynamic force in the country’s socio-political landscape.

*Druze* 

The Druze, a distinct religious group with roots in Islam, practice an esoteric faith that incorporates elements of Neoplatonism, Gnosticism, and other philosophical traditions. Known for their secrecy, the Druze do not proselytize, and their religious texts are accessible only to an elite group of initiates. Despite being a numerical minority, the Druze have maintained a significant political and social presence, often serving as power brokers in Lebanon's complex sectarian system. Politically, the Druze are represented through the leadership of influential families, most notably the Jumblatt family, who have played pivotal roles in national and regional politics. The Druze are concentrated in the Chouf Mountains, Mount Lebanon, and parts of the Bekaa Valley, where they form tight-knit communities that emphasize solidarity and mutual support. Their unique position in Lebanon's confessional mosaic allows them to navigate alliances flexibly, often bridging divides between larger religious groups while preserving their distinct identity and autonomy.

## III. Analysis

### 3.1 Spatial Analysis of Attacks

-   Present maps and tables showing:
    -   Attack frequencies by governorate.
    -   Refugee distributions relative to attack locations.
    -   Identity group distribution by governorate.

### 3.2 Likelihood of Attacks for given Identity

-   Calculate and present probabilities for each identity group using:
    -   P(attack \| identity) = n of attacks on groupTotal attacks in Lebanon
    -   P(attack \| identity)= Total attacks in LebanonNumber of attacks on group​
-   Discuss patterns or disparities.

```{r}
# Summarize attack data by governorate
attack_summary <- acled %>%
  group_by(adm1) %>%
  summarise(total_attacks = sum(events, na.rm = TRUE))

# Suppress grouping warning by using `.groups` argument
population_summary <- lebanon_confessional %>%
  group_by(adm1, sect) %>%
  summarise(
    total_population = sum(total_sect, na.rm = TRUE),  # Replace 'adm1_sect' with the actual population column name
    .groups = "drop"  # Remove grouping after summarise
  )

attack_summary <- acled %>%
  group_by(adm1) %>%
  summarise(
    total_attacks = sum(events, na.rm = TRUE),
    .groups = "drop"
  )

# Join population and attack summaries
combined_data <- population_summary %>%
  left_join(attack_summary, by = "adm1") %>%
  mutate(
    P_attack_given_identity = total_attacks / total_population
  )

# View the results
print(combined_data)

```

### 3.3 Casualties Per Identity Group

-   Show casualty distribution and normalize values by population size within each identity group for comparison.
-   Include visualizations such as bar charts or proportional symbol maps.

```{r}
normalized_casualties <- combined_data %>%
  group_by(sect) %>%
  summarise(
    total_population = sum(total_population, na.rm = TRUE),
    total_casualties = sum(P_attack_given_identity * total_population, na.rm = TRUE),  # Actual casualties
    normalized_casualties = total_casualties / total_population  # Normalize by population
  )

# View results
print(normalized_casualties)
```

```{r}
# Bar chart for normalized casualties
ggplot(normalized_casualties, aes(x = sect, y = normalized_casualties, fill = sect)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Normalized Casualties per Identity Group",
    x = "Identity Group",
    y = "Normalized Casualties (Casualties per Population)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r}
library(sf)  # For spatial data handling

# Combine normalized casualties with spatial geometries
map_data <- combined_data %>%
  mutate(total_casualties = P_attack_given_identity * total_population)  # Calculate casualties per region

# Proportional symbol map
ggplot(data = map_data) +
  geom_sf(aes(fill = total_casualties), color = NA) +
  scale_fill_viridis_c(option = "viridis", name = "Total Casualties") +
  labs(
    title = "Proportional Symbol Map of Total Casualties",
    caption = "Casualties derived from P_attack_given_identity."
  ) +
  theme_minimal()
```

### 3.4 Attack Types and Variations

-   Compare the types of attacks (e.g., airstrikes, ground operations) across identity groups. Use tables or clustered bar charts.
-   Highlight patterns or notable differences.

```{r}
# Summarize attack types by identity group
attack_type_summary <- combined_data %>%
  group_by(sect, event_type) %>%  # Group by identity and attack type
  summarise(
    count = n(),  # Count occurrences of each attack type
    .groups = "drop"
  )

# View the summary table
print(attack_type_summary)

```

```{r}
ggplot(attack_type_summary, aes(x = event_type, y = count, fill = sect)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Comparison of Attack Types Across Identity Groups",
    x = "Attack Type",
    y = "Number of Attacks",
    fill = "Identity Group"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Calculate proportions for each attack type within identity groups
attack_type_proportions <- attack_type_summary %>%
  group_by(sect) %>%
  mutate(
    proportion = count / sum(count, na.rm = TRUE)
  )

# View the proportion table
print(attack_type_proportions)

```

#### 3.5 Comparison of Attack Patterns Prior and Post October 7th (OPTIONAL)

-   Map and analyze attack frequencies and distributions of 1 year prior October 7th and post October 7th.
-   Perform statistical tests if applicable to assess whether attack rates differ significantly between time periods.

```{r descriptive_religion}
sect_summary <- confessional %>%
  group_by(sect) %>%
  summarise(sum = sum(total_sect)) %>%
  arrange(desc(sum))
sect_summary 
```
