---
title: "GIS Final Project"
author: "Homar A. Maurás Rodríguez"
format: html
editor: visual
execute: 
  echo: false
  error: false
---

```{r libraries}
library(tidyverse)
library(dplyr)
library(sf)
library(leaflet)
library(rnaturalearth)
library(mapview)
library(rnaturalearthhires)
```

# INTRODUCTION

## 2006 WAR ON LEBANON

## ESCALATIONS FOLLOWING OCTOBER 7TH
Include war declaration timeline

## OBJECTIVES

### RESEARCH QUESTIONS
1. What is P(attack | identity)P(attack | identity)?
2. Casualties per identity group.
3. Are types of attacks the same across identity groups?
4. Are Israeli attack rates across identity groups in Lebanon consistent between the 2006 Lebanon War and recent events?

### VARIABLES OF INTEREST
1. Identity
2. Location
3. Attack Sub-type
4. Time



## SIGNIFICANCE
Explain why understanding the spatial distribution of attacks across identity groups is important for humanitarian, political, and academic purposes.

## SCOPE
Define the geographic and demographic parameters (e.g., Lebanon’s 8 Governorates and key identity groups).


**METHOD OF ANALYSIS**
1. Time series analysis of 


**DATA DESCRIPTION**
This report pulls data from multiple sources to answer the proposed research questions:
1. Population estimates: U.S. State Department and WHO 
2. Israeli attacks on Lebanon: Using ACLED Data and filtering for IDF attacks
Administrative Lebanese Data:
Confessional Distribution Data:
2006 Lebanese War Reports:

```{r datasets}
#| error: false
acled_df <- read.csv("/Users/homi/Data Science/datasets/acled_aggregated_20048.csv")

unique(acled_df$Sub.Event.Type) #maybe we can organize it in a tibble to make it more presentable

#insert code for table

```

This is a map of Lebanon using mapview.

```{r lebanon_mapview}
lebanon_map <- st_read("/Users/homi/Data Science/datasets/geoBoundaries-LBN-ADM1-all/geoBoundaries-LBN-ADM1.shp")

# Combine Mount Lebanon and Jbiel

mapview(lebanon_map)
```

This is a map of Lebanon using leaflet \*prefered option

```{r leaflet}

```
